Deriving an Executable Model of a Data Stream Protocol from its Specification

### 2009-10-12 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: syntax errors corrections

There exist various approaches in literature addressing this kind of
development constraints (rapid development in hardware software
codesign), but they are all focused on big scale systems.  The
hardware software formal codesign methodologies usually have a common
structure\cite{Slomka2000,Hoffman2001}:

@@

Even when cosidering the development of the device hardware, embedded
software and software driver as separate entities that possibly have
to be developed concurrently, the approaches proposed are similar in
structure to the one seen in the case of hardware software
codesign\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
on very accurate formal specification that drive all the other
development tasks.

@@

To tackle the complexity of the concurrent tasks the formal model of
the protocol was used as a focus point that led all the other
development tasks (in this part, the approach is similar to the one
used in rapid development in hardware software codesign, as mentioned
above).  The protocol to specify has been divided in various (thin)
layers, each of the layer has been specified with a different
(lighweight) approach, some of the layers are completely formalized.
The formal model of the protocol was used to check at run time the
correctness of the /SB, the driver, the simulator, and the hand made
test suites of the driver and the simulator.  The formal model was
used to automatically generate a test suite to verify that both the
driver and the simulator were compliant to the specification, and
eventually to verify the specification itself.

@@

The generated test suite has a more focused and limited scope, since
only one method call is performed in each test, hence it is difficult
to explore complex behaviors. It is true that with a complete
specification, and specifying large enough data ranges (the
appropariate data ranges are combined to generate the tests), the
generated tests could theoretically achieve the same expressive power
as the hand made ones, but the price to pay would be expensive: much
longer and more complex specifications, requiring more time to be
designed and or updated, and more tests to be run, augmenting the
already not so short computation time of the generated test suite.

The test suites gave their best when used in a combined approach: the
generated test suite, checking most simple behaviors, while the hand
made checking the more complex ones.  During the development of the
driver code and the simulators, the generated test suit was helpful to
test the incrementally growing specification and the finer details of
each method, while the hand made one was helpful to verify the
behaviour on complex methods calls on objects intialized to specific
states. Having an hand made test suite also make the code style more
test oriented, which usually leads to a better quality and more
modularized source code\cite{Binder1999}.

@@

The combined test approach, driven by the packet protocol
specification, prooved succesfull on verifing the protocol itself and
the software implementing it.  Neverthless, it has to be noted that we
did not verified the highest levels of the protocol (see
\ref{sec:data_stream_and_protocol_description}: ``Packet sequence
rules'' and ``Packet input output asynchronous rules'').  Since on
these levels the protocol properties regards sequences of packets,
testing them can be tricky, since the sequences of packets are not
referred to explicitly in the implementation.  Two possible strategies
can be used:

\begin{itemize}
\item Translate the packet sequence properties in local packet
  properties (when this is possible).  The advantage is to have
  properties shown as the usual precondition, post conditions and
  invariants.  But the localized properties can be more complex and
  less intuitive and comprehensible than the original ones; finally,
  some of the global properties could proove themselves impossible to
  be translated into local ones.
\item Use separate utility classes, meant to verify a sequence of
  packages.  This approach is not clean, since it requires adding and
  using new classes, unneeded by the core implementation, to verify
  the protocol.  It has the advantage of keeping the properties that
  have to be verified simple.
\end{itemize}

One problem we noticed regards the executiuon time of the generated
tests.  The executiuon time is long for two reasons:

\begin{itemize}
\item Number of the tests generated. This problem is not directly
  addressable, since it is the heart of the test suite generation
  creation method itself: generating all the combinations of possible
  inputs for each method under test.
\item Runtime assertion checker. Each test is executed while RAC
  checks all the relavant properties. This problem, again, is not
  directly addressable. But, we can speculate that if the RAC
  assertions could be triggered on and off on single properties in a
  dynamic way, that is, during runtime, a test suite could use take
  advantage of this possibility, turning on the RAC only on the
  properties that have to be verified for the classes or methods under
  tests (and not for all the involved classes and methods); this could
  save time during the test suite execution.
\end{itemize}

@@

Approach:
\begin{itemize}
\item more specification used together: formal and not formal
\item formal specification is not complete: it is just good
  ``enough'', static checkers unfortunately cannot be used, since the
  specification is partial
\item use runtime asserion checker for formal specifications, on tests
\item use generated tests for formal specifications
\end{itemize}

Who drives what:
\begin{itemize}
\item specification (jml - formal specification or unit test -
  examples fragments) driven development
\item test driven specification
\item formal specification is not complete: it is just good
  ``enough'', static checkers unfortunately cannot be used, since the
  specification is partial
\item use runtime asserion checker for formal specifications, on tests
\item use generated tests for formal specifications
\end{itemize}

@@

Case example: actual paper, modified to fit the new parts, expecially
the test driven specification part.

>>>>

There exist various approaches in literature addressing this kind of
development constraints (rapid development in hardware software
co-design), but they are all focused on big scale systems.  The
hardware software formal co-design methodologies usually have a common
structure\cite{Slomka2000,Hoffman2001}:

@@

Even when considering the development of the device hardware, embedded
software and software driver as separate entities that possibly have
to be developed concurrently, the approaches proposed are similar in
structure to the one seen in the case of hardware software
co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
on very accurate formal specification that drive all the other
development tasks.

@@

To tackle the complexity of the concurrent tasks the formal model of
the protocol was used as a focus point that led all the other
development tasks (in this part, the approach is similar to the one
used in rapid development in hardware software co-design, as mentioned
above).  The protocol to specify has been divided in various (thin)
layers, each of the layer has been specified with a different
(lightweight) approach, some of the layers are completely formalized.
The formal model of the protocol was used to check at run time the
correctness of the /SB, the driver, the simulator, and the hand made
test suites of the driver and the simulator.  The formal model was
used to automatically generate a test suite to verify that both the
driver and the simulator were compliant to the specification, and
eventually to verify the specification itself.

@@

The generated test suite has a more focused and limited scope, since
only one method call is performed in each test, hence it is difficult
to explore complex behaviors. It is true that with a complete
specification, and specifying large enough data ranges (the
appropriate data ranges are combined to generate the tests), the
generated tests could theoretically achieve the same expressive power
as the hand made ones, but the price to pay would be expensive: much
longer and more complex specifications, requiring more time to be
designed and or updated, and more tests to be run, augmenting the
already not so short computation time of the generated test suite.

The test suites gave their best when used in a combined approach: the
generated test suite, checking most simple behaviors, while the hand
made checking the more complex ones.  During the development of the
driver code and the simulators, the generated test suit was helpful to
test the incrementally growing specification and the finer details of
each method, while the hand made one was helpful to verify the
behavior on complex methods calls on objects initialized to specific
states. Having an hand made test suite also make the code style more
test oriented, which usually leads to a better quality and more
modularized source code\cite{Binder1999}.

@@

The combined test approach, driven by the packet protocol
specification, proved successful on verifying the protocol itself and
the software implementing it.  Nevertheless, it has to be noted that we
did not verified the highest levels of the protocol (see
\ref{sec:data_stream_and_protocol_description}: ``Packet sequence
rules'' and ``Packet input output asynchronous rules'').  Since on
these levels the protocol properties regards sequences of packets,
testing them can be tricky, since the sequences of packets are not
referred to explicitly in the implementation.  Two possible strategies
can be used:

\begin{itemize}
\item Translate the packet sequence properties in local packet
  properties (when this is possible).  The advantage is to have
  properties shown as the usual precondition, post conditions and
  invariants.  But the localized properties can be more complex and
  less intuitive and comprehensible than the original ones; finally,
  some of the global properties could prove themselves impossible to
  be translated into local ones.
\item Use separate utility classes, meant to verify a sequence of
  packages.  This approach is not clean, since it requires adding and
  using new classes, unneeded by the core implementation, to verify
  the protocol.  It has the advantage of keeping the properties that
  have to be verified simple.
\end{itemize}

One problem we noticed regards the execution time of the generated
tests.  The execution time is long for two reasons:

\begin{itemize}
\item Number of the tests generated. This problem is not directly
  addressable, since it is the heart of the test suite generation
  creation method itself: generating all the combinations of possible
  inputs for each method under test.
\item Runtime assertion checker. Each test is executed while RAC
  checks all the relevant properties. This problem, again, is not
  directly addressable. But, we can speculate that if the RAC
  assertions could be triggered on and off on single properties in a
  dynamic way, that is, during runtime, a test suite could use take
  advantage of this possibility, turning on the RAC only on the
  properties that have to be verified for the classes or methods under
  tests (and not for all the involved classes and methods); this could
  save time during the test suite execution.
\end{itemize}

@@

Approach:
\begin{itemize}
\item more specification used together: formal and not formal
\item formal specification is not complete: it is just good
  ``enough'', static checkers unfortunately cannot be used, since the
  specification is partial
\item use runtime assertion checker for formal specifications, on tests
\item use generated tests for formal specifications
\end{itemize}

Who drives what:
\begin{itemize}
\item specification (jml - formal specification or unit test -
  examples fragments) driven development
\item test driven specification
\item formal specification is not complete: it is just good
  ``enough'', static checkers unfortunately cannot be used, since the
  specification is partial
\item use runtime assertion checker for formal specifications, on tests
\item use generated tests for formal specifications
\end{itemize}

@@

Case example: actual paper, modified to fit the new parts, especially
the test driven specification part.
